Maintenance planning on French military aircraft operations

### 2017-11-27 latex/ROADEF/article_roadef_2018.tex
experiments with model and changes after Olga's revision

\section{Introduction}

    The Military Flight and Maintenance Planning is a problem consisting on the assignment of missions and maintenances tasks to military aircraft. It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically but they are assumed to include a round-trip to the base. It also includes different objectives and constraints.

    Although the former has been studied in literature in \cite{Cho2011}, \cite{Chastellux2017}, \cite{Kozanidis2008}, it has not received as much attention as the latter.

@@

\section{The problem}

    The problem consists in the planning of the maintenance operations for military aircraft while complying with already scheduled missions. In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks".

    % \subsection{Tasks}
    % \label{def:task}

    A series of $v \in \mathcal{V}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Each task require a certain number $rr_{v}$ of resources $a \in \mathcal{A}$ and, while assigned, it wears them for a quantity of time defined by $rh_{v}$ in each period. There is a set $cd_{v} \subset \mathcal{A}$ that lists the resources that can be assigned to each task.

@@

    Resources suffer from wear and tear and so require a recurrent preventive maintenance in order to be in a correct state. These maintenances need to take place before the first of two things occurs: a given amount of elapsed time $me$ has passed since the previous maintenance or a given amount of usage time $mu$ has passed since the previous maintenance. The following are the possible $s \in \mathcal{S}$ states for a resource to be in any given time: Assigned to a task ('V'), Receiving maintenance ('M'), Available ('A'), Needing maintenance ('N'). These maintenances take an amount of time equal to $md$ periods and cannot be interrupted. The number of resources under maintenance in any given period is limited by $c_t$.

    % We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance at any given period.

@@

\section{The model}

@@

    The decisions involve, in essence, the management of the previously defined resources. We can summarize the main decisions the following way: if resource $a$ is in state $s$ in each period $t$ and if resource $a$ is assigned to task $v$ in period $t$.

    \begin{tabular}{p{15mm}lp{105mm}}
        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assigned task $v$ in period $t$, 0 otherwise. \\  
        $state_{ast}$       & : & 1 if resource $a$ is assigned state $s$ in period $t$, 0 otherwise. \\  
        $start_{at}$        & : & 1 if resource $a$ starts maintenance in period $t$, 0 otherwise. \\
        $ret_{at}$          & : & remaining elapsed time (integer) for resource $a$ after period $t$. \\  
        $rut_{at}$          & : & remaining usage time (integer) for resource $a$ after period $t$. \\  
        $used_{at}$         & : & time (integer) used by resource $a$ in period $t$. \\
        $min\_avail$        & : & min number (integer) of available resources in any period. \\
        $max\_maint$        & : & max number (integer) of resources in maintenance in any period. \\
    \end{tabular}
    \bigskip

@@

    The objective is to both minimize the maximum number of maintenances $\text{Min}\, (max\_maint)$ and maximize the minimal number of available aircraft \text{Max}\, (min\_avail) subject to the constraints listed below.

    % \begin{align}
    %     & \text{Min}\, (max\_maint) \\
    %     & \text{Max}\, (min\_avail)
    % \end{align}

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % MAIN CONSTRAINTS
        %%%%%%%%%%%%%%%%%%%
        % maximum capacity:
        & \sum_{a \in \mathcal{A}} state_{ast} \leq c_{t}
                & t \in \mathcal{T}, s = \mq{M} \label{eq:capacity}\\
        % minimum resources
        & \sum_{a \in cd_v} task_{avt} \geq rr_{t}
                & t \in \mathcal{T}, v \in \mathcal{V},  \label{eq:taskres}\\
        % max one task per period
        & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
        % if assigned to task: in working state
        & state_{ast} \geq task_{avt} 
                & \hspace{10mm} t \in \mathcal{T}, a \in cd_v, v \in \mathcal{V}, s = \mq{V} \label{eq:state1}\\
        % if no working state, no assigned to any task
        & state_{ast} + task_{avt} \leq 1
                & \hspace{10mm} t \in \mathcal{T}, a \in cd_v, v \in \mathcal{V}, s \neq \mq{V} \label{eq:state3}\\
        % only one state. If we take out the 'do nothing' state, this could be \leq
        & \sum_{s \in \mathcal{S}} state_{ast} = 1 
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}
    \end{align}

@@

\section{Conclusions}

    The present modelization is a new way to formulate the Military Flight Planning Model. Unlike previous work, it assigns aircraft to missions instead of just using an aggregate hourly demand per period. This is useful in order to improve the model by including minimum durations for mission assignment.

>>>>

\section{Introduction}
    The Military Flight and Maintenance Planning is a problem consisting on the assignment of missions and maintenance tasks to military aircraft. It is a variant of the Flight and Maintenance Planning problem where flights are not modeled geographically but they are assumed to include a round-trip to the base. It also includes different objectives and constraints. Although the former has been studied in literature in \cite{Cho2011,Chastellux2017,Kozanidis2008}, it has not received as much attention as the latter.

@@

%\section{The problem}

    The problem consists in the planning of the maintenance operations for military aircraft and the assignment of these aircraft to already scheduled missions. In order to generalize the formulation, from now on aircraft will be named "resources" and missions will be named "tasks". A series of $v \in \mathcal{V}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Each task requires a certain number $rr_{v}$ of resources $a \in \mathcal{A}$ which it employs for a time duration defined by $rh_{v}$ in each period. Set $cd_{v} \subset \mathcal{A}$ lists the resources that can be assigned to each task.

@@

    Resources require recurrent preventive maintenance operations. The frequency of maintenance operations is defined by 2 factors: the amount of time $me$ elapsed since the last maintenance operation and the amount of time $mu$ for which the resource has been used since the last maintenance operation. The following states $s \in \mathcal{S}$ are possible for a resource: assigned to a task ('V'), receiving maintenance ('M'), available ('A'), needing maintenance ('N'). A maintenance operation takes an amount of time equal to $md$ periods and cannot be interrupted. The number of resources under maintenance in any period is restricted by its capacity $c_t$. In the following we present a new mathematical model for this problem. Unlike previous work, it assigns aircraft to missions, instead of just using an aggregate demand. 
    % We will call "remaining elapsed time" (or $ret$) to the amount of time that needs to pass in order for the resource to need a maintenance at any given period.

@@

\section{Mathematical model}

@@

    The decision variables manage the previously defined resources. We can summarize the main decisions the following way: if resource $a$ is in state $s$ in each period $t$ and if resource $a$ is assigned to task $v$ in period $t$.

    \begin{tabular}{p{15mm}lp{105mm}}
        $task_{avt}$        & : & 1 if resource $a \in cd_v$ is assigned to task $v$ in period $t$, 0 otherwise. \\  
        $state_{ast}$       & : & 1 if resource $a$ is in state $s$ in period $t$, 0 otherwise. \\  
        $start_{at}$        & : & 1 if resource $a$ starts maintenance in period $t$, 0 otherwise. \\
        $ret_{at}$          & : & remaining elapsed time (integer) for resource $a$ after period $t$. \\  
        $rut_{at}$          & : & remaining usage time (integer) for resource $a$ after period $t$. \\  
        $used_{at}$         & : & time (integer) used by resource $a$ in period $t$. \\
        $min\_avail$        & : & min number (integer) of available resources in any period. \\
        $max\_maint$        & : & max number (integer) of resources in maintenance in any period. \\
    \end{tabular}
    \bigskip

@@

    The objective is to both minimize the maximum number of maintenances and maximize the minimal number of available aircraft:

    \begin{align}
        & \text{Min}\, (max\_maint) \\
        & \text{Max}\, (min\_avail)
    \end{align}

    Subject to:

    \begin{align}
        %%%%%%%%%%%%%%%%%%%
        % MAIN CONSTRAINTS
        %%%%%%%%%%%%%%%%%%%
        % maximum capacity:
        & \sum_{a \in \mathcal{A}} state_{ast} \leq c_{t}
                & t \in \mathcal{T}, s = \mq{M} \label{eq:capacity}\\
        % minimum resources
        & \sum_{a \in cd_v} task_{avt} \geq rr_{v}
                & t \in \mathcal{T}, v \in \mathcal{V},  \label{eq:taskres}\\
        % max one task per period
        & \sum_{v \in \mathcal{V}} task_{avt} \leq 1
                & t \in \mathcal{T}, a \in \mathcal{A} \label{eq:maxtasks}\\
        % if assigned to task: in working state
        & state_{ast} \geq task_{avt} 
                & \hspace{10mm} t \in \mathcal{T}, a \in cd_v, v \in \mathcal{V}, s = \mq{V} \label{eq:state1}\\
        % if no working state, no assigned to any task
        & state_{ast} + task_{avt} \leq 1
                & \hspace{10mm} t \in \mathcal{T}, a \in cd_v, v \in \mathcal{V}, s \neq \mq{V} \label{eq:state3}\\
        % only one state. If we take out the 'do nothing' state, this could be \leq
        & \sum_{s \in \mathcal{S}} state_{ast} = 1 
                & \hspace{10mm} t \in \mathcal{T}, a \in \mathcal{A} \label{eq:state2}
    \end{align}

@@

%\section{Conclusions}

 %   The present modelization is a new way to formulate the Military Flight Planning Model. Unlike previous work, it assigns aircraft to missions instead of just using an aggregate hourly demand per period. This is useful in order to improve the model by including minimum durations for mission assignment.
